Ensemble Methods for Robust Feature Selection

### 2016-08-05 paper/paper.tex
small changes

The second feature selection is RELIEFF \citep{kononenko1997overcoming}. For each sample $x_i$ the k nearest neighboring samples, using the $L_1$ norm, belonging the same class $\{\textrm{Near-hit}_{i, j}\}_{j=1..k}$ and the ones of the opposite class $\{\textrm{Near-miss}_{i, j}\}_{j=1..k}$ are selected. Then the difference is taken to update the weight vector representing how well each feature distinguishes samples of the two classes, as can be seen in equation \ref{eq:relief}. In all the experiments k is set to 5, which, empirically, gives satisfactory results.
\begin{equation}
  \label{eq:relief}
  W_i = W_i - \frac{1}{k}\sum_{j=1}^{k}|x_i - \textrm{Near-hit}_{i,j}| + \frac{1}{k}\sum_{j=1}^{k} |x_i - \textrm{Near-miss}_{i,j}|
\end{equation}

The third feature selection method used is a recursive feature elimination (RFE) with a C-SVM classifier \citep{guyon2002gene}. 
The recursive feature elimination fits a classifier and ranks the feature according to their importance in the model. In the case of the C-SVM, it is based on the weight vector of the hyperplane. More precisely after obtaining the solution for the primal problem of the C-SVM in equation \ref{eq:svm_primal}, the ranking coefficients $c_i$ for the features are then calculated by squaring the weights $c_i = w_i^2$.
\begin{alignat}{-1}
     \min_{w,b}  & \quad \frac12\|w\|^2 + C\sum_{i=1}^m \xi_i & & & \label{eq:svm_primal} \\ 
   \text{s.\,t.} & \quad y^{(i)} (w^Tx^{(i)} + b) & \quad \geq & \quad 1 -\xi_i &&
                   \quad \forall \, i = 1, \ldots, m \nonumber \\
                 & \quad \xi_i                  & \quad \geq & \quad 0 &&
                   \quad \forall \, i = 1, \ldots, m \nonumber 
\end{alignat}

The worst ranked features are then discarded and the procedure is repeated until only a subset of features is left. In the experiments, $10\%$ of the size of all features are discarded at each step. The procedure is repeated until a subset of the size of $1\%$ of all features is left. Since each step $10\%$ of the features are discarded, the overall process takes $10$ steps. Therefore a C-SVM is $10$ times needed to be evaluated. Another method would be to discard only $10\%$ of the remaining features. A more gradually ascend of feature weights can be achieved this way. However, it is not used due to the substantially increase in needed computation power. For the hpyerparameter $C$ a grid search on the set $\{1, 10, 100, 1000\}$ is done. The parameters are evaluated with stratified 5-fold cross-validation. In this work the RFE method with a C-SVM classifier is called SVM RFE.

The last feature selection method used is Lasso logistic regression \citep{wheeler2010lasso}. It is a logistic regression with a $L_1$ norm as penalty for the weight vector. The hyperparameter $C$ is estimated with a grid search with $10$ different parameters in logarithmic scale between $10^{-5}$ and $10^3$. The parameters are evaluated with a stratified 3-fold cross-validation.

\begin{equation}
    \min_{w,c} \quad \|w\|_1 + C\sum_{i=1}^n \log(\exp(y_i(x_i^Tw + c))+1)
  \end{equation}
  The advantage of this method is to obtain a sparse weight vector as solution. Hence only a small subset of features have a significant high weight. In this work Lasso logistic regression is also called Lasso for simplicity.

@@

\begin{description}[align=left]
\item [Linear aggregation :] The average of the weights is taken as the new weight. Two version are used, one where a normalization of the weight sum is applied and the other without it.

@@

\subsection{Data Sets}
Datasets are taken from the NIPS 2003 challenge (\cite{NIPS}). However, due to computation power constraints, not the whole dataset is always used. Additionally the colon dataset (\cite{alon1999broad}) is used. Finally, artificial data is also generated, since here it is possible to control which features are meaningful.

@@

\subsection{Results}
In this section the results of the different experiments are presented. First the distribution of the weights given by the feature selectors is analyzed. Secondly the results on the artificial dataset is discussed. Then the influence of different combination of feature selectors is analyzed. Finally the a comparison between the ensemble methods and the feature selectors is made.

>>>>

The second feature selection is RELIEFF \citep{kononenko1997overcoming}. For each sample $x_i$ the k nearest neighboring samples, using the $L_1$ norm as metric, belonging to the same class $\{\textrm{Near-hit}_{i, j}\}_{j=1..k}$ and the ones of the opposite class $\{\textrm{Near-miss}_{i, j}\}_{j=1..k}$ are selected. Then the difference is taken to update the weight vector representing how well each feature distinguishes samples of the two classes, as can be seen in equation \ref{eq:relief}. In all the experiments k is set to 5, which empirically gives satisfactory results.
\begin{equation}
  \label{eq:relief}
  W_i = W_i - \frac{1}{k}\sum_{j=1}^{k}|x_i - \textrm{Near-hit}_{i,j}| + \frac{1}{k}\sum_{j=1}^{k} |x_i - \textrm{Near-miss}_{i,j}|
\end{equation}
The third feature selection method uses recursive feature elimination (RFE) with a linear C-SVM classifier \citep{guyon2002gene} to weight the features. 
The RFE method fits a classifier and ranks the feature according to their importance in the model. In the case of the C-SVM, it is based on the weight vector of the hyperplane. More precisely after obtaining the solution for the primal problem of the C-SVM in equation \ref{eq:svm_primal}, the ranking coefficients $c_i$ for the features are then calculated by squaring the weights $c_i = w_i^2$.
\begin{alignat}{-1}
     \min_{w,b}  & \quad \frac12\|w\|^2 + C\sum_{i=1}^m \xi_i & & & \label{eq:svm_primal} \\ 
   \text{s.\,t.} & \quad y^{(i)} (w^Tx^{(i)} + b) & \quad \geq & \quad 1 -\xi_i &&
                   \quad \forall \, i = 1, \ldots, m \nonumber \\
                 & \quad \xi_i                  & \quad \geq & \quad 0 &&
                   \quad \forall \, i = 1, \ldots, m \nonumber 
\end{alignat}
The worst ranked features are then discarded and the procedure is repeated until only a subset of features is left. In the experiments, $10\%$ of the size of all features are discarded at each step. The procedure is repeated until a subset of the size of $1\%$ of all features is left. Since in each step $10\%$ of all features are discarded, the overall process takes $10$ steps. Therefore a C-SVM is needed to be $10$ times evaluated. Another method would be to discard only $10\%$ of the remaining features. A more gradually ascend of feature weights can be achieved this way. However, it is not used due to the substantially increase in needed computation power. For the hpyerparameter $C$ a grid search on the set $\{1, 10, 100, 1000\}$ is done. The parameters are evaluated with stratified 5-fold cross-validation. In this work the RFE method with a C-SVM classifier is abbreviated as SVM RFE.

The last feature selection method used is Lasso logistic regression \citep{wheeler2010lasso}. It is a logistic regression with a $L_1$ norm as penalty for the weight vector. The hyperparameter $C$ is estimated with a grid search with $10$ different parameters in logarithmic scale between $10^{-5}$ and $10^3$. The parameters are evaluated with a stratified 3-fold cross-validation.
\begin{equation}
    \min_{w,c} \quad \|w\|_1 + C\sum_{i=1}^n \log(\exp(y_i(x_i^Tw + c))+1)
  \end{equation}
  The advantage of this method is to obtain a sparse weight vector as solution. Hence only a small subset of features have a significant high weight. In this work Lasso logistic regression is also abbreviated as Lasso for simplicity.

@@

\begin{description}[align=left]
\item [Linear aggregation :] The average of the weights is taken as the new weight. Two version are used, one where a normalization of the weight sum is applied and one without this normalization.

@@

\subsection{Data Sets}
Datasets are taken from the NIPS 2003 challenge (\cite{NIPS}). However, due to computation power constraints, not the whole dataset is always used. Additionally the colon dataset (\cite{alon1999broad}) is used. Finally, artificial data is also generated, since here it is possible to control which features are meaningful. All data sets are binary labelled.

@@

\subsection{Results}
In this section the results of the different experiments are presented. Firstly, the distribution of the weights given by the feature selectors is analyzed. Secondly, the results on the artificial dataset is discussed. Afterwards the influence of different combinations of feature selectors is analyzed. Finally, a comparison between the ensemble methods and the feature selectors is made.
